---
title: "Testing BACON on Simulated Datasets"
author: "Kevin Jin"
date: "2023-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Generate simulated shape data via simulation of polygonal chains.
Step 0: Load essential functions for simulating shape data.
```{r}
source("~/Documents/Repositories/BACON/code/data_simulation/functions.R")
```

Step 1: Generate a dataset containing 10 clusters of a total of 1000 3-gons, 
each cluster having 100 3-gons, each with 0.02 jitter.

Dataset structure:
> 1. Cluster 1
> + Shape 1
> + Shape 2
> ...
> + Shape 100
> 2. Cluster 2
> ...
> 3. Cluster 10
> ...

```{r}
x <- 200 # Total number of n-gon
z <- 10 # Number of clusters
n <- (x / z) # Number of n-gons per cluster
k <- 5 # Number of vertices in each n-gon
jitter_factor <- 0.02

dataset <- simulate_shapes(x, z, n, k, jitter_factor)
```

Step 2: For a sanity check, plot each shape to visualize it.
```{r echo=FALSE, eval = FALSE}
for (i in seq_along(dataset)) {
  for (shape in dataset[[i]]) {
    plot(shape, type = "l")
  }
}
```

Step 3: Extract side length and angle proportions as the data to be clustered. 
Creates nested lists for angle proportions and side length proportions.
```{r}
angles <- matrix(nrow = x, ncol = k, byrow = TRUE)
side_lengths <- matrix(nrow = x, ncol = k, byrow = TRUE)

counter <- 1
for (i in seq_along(dataset)) {
  for (j in dataset[[i]]) {
    angles[counter, ] <- get_interior_angles(j)
    side_lengths[counter, ] <- get_side_lengths(j)
    counter <- counter + 1
  }
}

# Clean up variables
rm(i, j)
```

Step 4: Test the BACON clustering algorithm.
```{r}
# Source the MCMC algorithm
setwd("~/Documents/Repositories/BACON/code/clustering/")
source("bacon.R")
```

```{r eval = FALSE}
# Cluster dataset with BACON
res <- bacon(side_lengths, 
             angles, 
             K = z,
             weight_L = 1,
             weight_A = 1,
             estimate.s = TRUE, 
             estimate.r = TRUE,
             iter = 10000, 
             burn = 5000)
res$cluster # BACON-estimated cluster memberships
res$s_map # BACON-estimated shift parameters for shape registration
res$r_map # BACON-estimated reverse parameters for shape registration

## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
ground_truth <- rep(1:z, each = n)
mclust::adjustedRandIndex(res$cluster, ground_truth)
```

Step 5a: Test the effect of side length weight on model performance.
```{r echo = FALSE, eval = FALSE}
weight_performance <- as.data.frame(matrix(nrow = 10, ncol = 2, byrow = TRUE))
colnames(weight_performance) <- c("weight", "ari")
ground_truth <- rep(1:z, each = n)
  
row <- 1
for (w in seq(from = 0, to = 1, by = 0.1)) {
  res <- bacon(side_lengths, 
               angles, 
               K = z,
               weight_L = w,
               weight_A = 1,
               estimate.s = FALSE, 
               estimate.r = FALSE,
               iter = 10000, 
               burn = 5000)
  weight_performance[row, "weight"] <- w
  weight_performance[row, "ari"] <- mclust::adjustedRandIndex(res$cluster, 
                                                         ground_truth)
  row <- row + 1
}

plot(weight_performance,
     xlab = "Side Length Weight", 
     ylab = "ARI",
     main = "BACON Performance based on Side Length Weight (200 3-gons)",
     type = "b")
```

Step 5b: Test the effects of both angle weight and side length weight on model
performance.
```{r echo = FALSE}
ground_truth <- rep(1:z, each = n)
n_replicates <- 10
weight_performances <- vector("list", n_replicates)

# Run every simulation setting 5-10 times and take the mean ARI
for(i in 1:length(weight_performances)) {
  # Create matrix to store ARI results
  weight_performance <- matrix(nrow = 5, ncol = 5, byrow = TRUE)
  colnames(weight_performance) <- seq(from = 0, to = 2, by = 0.5)
  rownames(weight_performance) <- seq(from = 0, to = 2, by = 0.5)
  # Run BACON across weights [0, 2] and calculate ARI
  row <- 1
  col <- 1
  for (l in seq(from = 0, to = 2, by = 0.5)) {
    for (a in seq(from = 0, to = 2, by = 0.5)) {
      res <- bacon(side_lengths, 
                   angles, 
                   K = z,
                   weight_L = l,
                   weight_A = a,
                   estimate.s = FALSE, 
                   estimate.r = FALSE,
                   iter = 5000, 
                   burn = 2500)
      weight_performance[row, col] <- mclust::adjustedRandIndex(res$cluster, ground_truth)
      col <- col + 1
    }
    row <- row + 1
    col <- 1
  }
  # Store matrix in master list
  weight_performances[[i]] <- weight_performance
}

# Take the element-wise mean of all weight performances
weight_performance <- Reduce("+", weight_performances) / length(weight_performances)

require(pheatmap)
require(grid)
setHook("grid.newpage", function() pushViewport(viewport(x=1,y=1,width=0.94, height=0.94, name="vp", just=c("right","top"))), action="prepend")
pheatmap(mat = weight_performance,
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         display_numbers = TRUE,
         main = "Mean BACON performance over 10 replicates (n = 200; k = 5; z = 10)",
         xlab = "Angle Weight", 
         ylab = "Side Length Weight",
         scale = "none")
setHook("grid.newpage", NULL, "replace")
grid.text("Angle Weight", y=-0.03, gp=gpar(fontsize = 12))
grid.text("Side Length Weight", x=-0.03, rot=90, gp=gpar(fontsize = 12))
```
Step 6: Reconstruct all closed unit polygonal chains for each shape from the 
side length and angle proportions. 

**WARNING: Computationally intensive!**
Number of possible shapes calculated increases exponentially with the number
of vertices per shape and the number of shapes in the dataset.
```{r}
pc <- reconstruct(angles[1, ], side_lengths[1, ])
plot(data.frame(pc[[1]]), type = "l")

for (row in 1:nrow(angles)) {
  pc <- reconstruct(angles[row, ], side_lengths[row, ])
  #plot(data.frame(pc[[1]]), type = "l")
  #plot(data.frame(pc[[2]]), type = "l")
}
```

# Generate simulated data by drawing from a truncated Dirichlet distribution.

Step 0: Load essential functions for simulating TDD data.
```{r}
source("~/Documents/Repositories/BACON/code/data_simulation/functions.R")
```

Step 1: Simulate side length and angle proportions for 17 3-gons (Reverse
mapping function seems to add an extra vertex? Is it because it's not closed?):
```{r}
# Generate side length proportions for a dataset of 3-gons
z1_length <- 10
z2_length <- 5
z3_length <- 2
z4_length <- 29
z1 <- rtdirichlet(n = z1_length, eta = c(1, 1, 1), a = c(0, 0, 0), b = c(0.5, 0.5, 0.5))
z2 <- rtdirichlet(n = z2_length, eta = c(2, 1, 2), a = c(0, 0, 0), b = c(0.5, 0.5, 0.5))
z3 <- rtdirichlet(n = z3_length, eta = c(5, 2, 1), a = c(0, 0, 0), b = c(0.5, 0.5, 0.5))
z4 <- rtdirichlet(n = z4_length, eta = c(3, 5, 4), a = c(0, 0, 0), b = c(0.5, 0.5, 0.5))
side_lengths <- rbind(z1, z2, z3, z4)

# Generate angle proportions for a dataset of 3-gons
z1_length <- 10
z2_length <- 5
z3_length <- 2
z4_length <- 29
z1 <- rtdirichlet(n = z1_length, eta = c(1, 1, 1), a = c(0, 0, 0), b = c(0.3, 0.3, 0.3))
z2 <- rtdirichlet(n = z2_length, eta = c(2, 1, 2), a = c(0, 0, 0), b = c(0.3, 0.3, 0.3))
z3 <- rtdirichlet(n = z3_length, eta = c(5, 2, 1), a = c(0, 0, 0), b = c(0.3, 0.3, 0.3))
z4 <- rtdirichlet(n = z4_length, eta = c(3, 5, 4), a = c(0, 0, 0), b = c(0.3, 0.3, 0.3))
angles <- rbind(z1, z2, z3, z4)

# Generate ground truth
ground_truth <- c(rep(1, z1_length), rep(2, z2_length), rep(3, z3_length), rep(4, z4_length))
```

Step 2: Test BACON clustering algorithm.
```{r}
setwd("~/Documents/Repositories/BACON/code/clustering/")
source("bacon.R")
```

```{r}
res <- bacon(side_lengths, 
             angles, 
             K = 4, 
             weight_L = 1,
             weight_A = 1,
             estimate.s = FALSE, 
             estimate.r = FALSE,
             iter = 10000, 
             burn = 5000)
res$cluster
res$s_map
res$r_map

## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
mclust::adjustedRandIndex(res$cluster, ground_truth)
```

Step 3: Reconstruct all closed unit polygonal chains for each shape from the 
side length and angle proportions. 
```{r}
for (i in 1:nrow(angles)) {
  assign(paste0("pc_", i), reconstruct(angles[i, ], side_lengths[i, ]))
}
```
# Run BACON on Huimin's simulated demo data.
```{r}
## The following guide demonstrates clustering of exmaple data using BACON.
## BACON requires three inputs:
## 1. L: a m-by-n matrix of length proportions, where m is the number of n-gons 
## and n is the number of gons.
## 2. A: a m-by-n matrix of angle proportions.
## 3. K: the number of clusters in the dataset.
```

Step 0: Load required packages.
```{r}
## Load required packages
setwd("~/Documents/Repositories/BACON/code/clustering")
source("bacon.R")
```

Step 1: Load Huimin's simulated data.
```{r}
## Load demo data (100 triangles with 3 clusters)
setwd("~/Documents/Repositories/BACON/data/simulated/Data")
load("demo.RData")
head(L)
head(A)
```

Step 2: Run BACON on the data.
```{r}
## Run the model (Default iterations = 2000; burn in = 1000)
res <- bacon(L, A, K = 3, iter = 10000, burn = 5000) # ARI = 0.536 ~ 0.731
res <- bacon(L2, A2, K = 3) # ARI = 1
```

Step 3: Check convergence and clustering accuracy.
```{r}
## BACON produces the following output variables:
## cluster - the estimated cluster assignment.
## s_map - the estimated starting vertex indicators. 
## r_map - the estimated reverse indicators r.
res$cluster
res$s_map
res$r_map

## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
mclust::adjustedRandIndex(res$cluster, z)
```
# Run BACON on "Decagons" simulated data.

Step 0: Load required packages.
```{r}
#### Run BACON on simulated 20-gon ("decagons") data. ####
## Load required packages
setwd("~/Documents/Repositories/BACON/code/clustering")
source("bacon.R")
```

Step 1: Load simulated 20-gons data.
```{r}
## Load simulated shape data (100 20-gons with 10 clusters)
setwd("~/Documents/Repositories/BACON/data/simulated/Data")
load("decagons.Rdata")
head(side_lengths)
head(angles)
```

Step 2: Run the model.
```{r}
## Run the model
res <- bacon(side_lengths[, 1:20], 
             angles[, 1:20], 
             K = 10, 
             weight_A = 1, 
             weight_L = 1,
             estimate.s = TRUE, 
             estimate.r = TRUE,
             iter = 10000, 
             burn = 5000)
```

Step 3: Check convergence and clustering accuracy.
```{r}
## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
mclust::adjustedRandIndex(res$cluster, angles[, 21]) 

# ARI: 0.46 at 10000/5000 iter
```