---
title: "Testing BACON on Real Datasets"
author: "Kevin Jin"
date: "2023-05-16"
output: html_document
---

# Run BACON on polygonal chains converted from MPEG-7.
Step 0: Load required packages.
```{r}
#### Run BACON on MPEG-7 data. ####
## Load required packages
setwd("~/Documents/Repositories/BACON/code/clustering")
source("bacon.R")
```

Step 1: Load MPEG-7 polygonal chains.
```{r}
## Load MPEG-7 data
setwd("~/Documents/Repositories/BACON/data/real/MPEG-7")
load("MPEG-7_k=20.Rdata")
head(side_lengths)
head(angles)
```

Step 2: Run the model.
```{r}
## Run the model
res <- bacon(side_lengths, 
             angles, 
             K = 20,
             weight_L = 1,
             weight_A = 1,
             estimate.s = FALSE, 
             estimate.r = FALSE,
             iter = 10000, 
             burn = 5000)
```

Step 3: Check convergence and clustering accuracy.
```{r}
## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
mclust::adjustedRandIndex(res$cluster, phenotype)
```

# Run BACON on polygonal chains converted from Animals (Curated MPEG-7).

# Run BACON on polygonal chains converted from ADHD-200.
Step 0: Load required packages.
```{r}
#### Run BACON on ADHD-200 data. ####
## Load required packages
setwd("~/Documents/Repositories/BACON/code/clustering")
source("bacon.R")
```

Step 1: Load ADHD-200 data.
```{r}
## Load ADHD-200 data (647 50-gons with 4 clusters)
setwd("~/Documents/Repositories/BACON/data/real/adhd_200")
load("ADHD-200_compositional.Rdata")
head(side_lengths)
head(angles)
```

Step 1a: Reduce ADHD-200 vertex number using BayesLASA (?)
```{r}
## Reduce number of vertices to speed up computation
# Apply BayesLASA to reduce 50-gons to 10-gons - also for other datasets like MPEG-7 and ETH-80 (?)
# What is the minimum number of vertices needed to describe any object in this dataset? It may vary.
```

Step 2: Run the model.
```{r}
## Run the model
res <- bacon(side_lengths[, 1:50], 
             angles[, 1:50], 
             K = 4,
             weight_L = 1,
             weight_A = 1,
             estimate.s = TRUE,
             estimate.r = TRUE,
             iter = 10000, 
             burn = 5000) 
```

Step 3: Check convergence and clustering accuracy.
```{r}
## Check convergence
plot(rowSums(res$s_store == 0), type = "l", 
     ylab = "Number of samples with starting point as 0", xlab = "Iteration")

## Check clustering accuracy
mclust::adjustedRandIndex(res$cluster, phenotype)

# ARI: 0.027; User runtime: 95 minutes
# ARI: 0.022 at 10000/5000 iter; Elapsed runtime: 104 minutes
```